Benefits and risks associated with children’s and adolescents’ interactions with electronic screens: An umbrella review

### 2023-03-10 reports/manuscript.Rmd
Some small tweaks

affiliation:
  - id            : "1"
    institution   : "Institute for Positive Psychology and Education, Australian Catholic University"
  - id            : "2"
    institution   : "School of Health and Behavioural Sciences, Australian Catholic University"
  - id            : "3"
    institution   : "Department of Sport Science and Clinical Biomechanics, University of Southern Denmark"
  - id            : "4"
    institution   : "Centre for Health Research, University of Southern Queensland"
  - id            : "5"
    institution   : "Department of Health Systems and Populations, Faculty of Medicine, Health and Human Sciences, Macquarie University"
  - id            : "6"
    institution   : "School of Kinesiology, Louisiana State University"
  - id            : "7"
    institution   : "School of Medicine and Health, Sydney University"
  - id            : "8"
    institution   : "Faculty of Health Sciences, Curtin University"
  - id            : "9"
    institution   : "Department of Movement and Sport Science, Ghent University"
  - id            : "10"
    institution   : "Institute for Physical Activity and Nutrition, Deakin University"
  - id            : "11"
    institution   : "The Matilda Centre for Research in Mental Health and Substance Use, University of Sydney"
  - id            : "12"
    institution   : "School of Education, Australian Catholic University"

@@

documentclass     : "apa6"
classoption       : "man"
csl               : nature.csl
output            : papaja::apa6_pdf
header-includes:
    - \usepackage{makecell}
    - \usepackage{booktabs}
    - \usepackage{colortbl}
---

@@

To facilitate comparisons, we converted effect sizes to Pearson's $r$ using established formulae [@bowmanEffectSizesStatistical2012;@jacobsEstimationBiserialCorrelation2017].
Effect sizes on the original metric are provided in Supplementary File 2.</span>
<span id ="r2_27">Throughout the results section we interpret the size of the effects using Funder and Ozer's guidelines [@funderEvaluatingEffectSize2019]: very small (0.05 < r <= 0.1), small (0.1 < r <= 0.2), medium (0.2 < r <= 0.2), large (0.3 < r <= 0.4), and very large (r >= 0.4).
These are similar to other interpretations based on empircal data [@gignacEffectSizeGuidelines2016].</span>

@@

The most frequently reported exposures were `r manuscript_info$freq_exposures`.
Supplementary File 4 provides a list of all exposures identified.
The most frequently reported outcomes were `r manuscript_info$freq_outcomes`.
<span id ="r2_25">Of the `r manuscript_info$unique_combos` unique exposure/outcome combinations, `r manuscript_info$freq_combos$n1` occured in only one review, with `r manuscript_info$freq_combos$n2` appearing twice, and `r manuscript_info$freq_combos$ngt2` appearing three or more times.
Full characteristics of the included studies are provided in Table 1.</span>
<span id ="r2_23">After removing reviews with duplicate exposure/outcome combinations, our process yielded `r manuscript_info$unique_effects` unique effect/outcome combinations (retaining multiple effects for different age groups or study designs) contributed from `r manuscript_info$prisma$unique_reviews` reviews.</span>
These effects represent the findings of `r format(manuscript_info$effects_sum$sum_k, big.mark=",")` primary studies comprised of `r format(manuscript_info$effects_sum$sum_n, big.mark=",")` participants.

@@

Screen use research is extensive, varied, and rapidly growing.
Reviews tended to be general (e.g., all screen time) and even when more targeted (e.g., social media) nuances related to specific content (e.g., Instagram vs Facebook) have not been meta-analysed or have not produced credible evidence.
<span id = "r2_29">Fewer than 20\% of the effects identified met our criteria for statistical credibility.
Most studies which did not meet our critiera failed to provide study-level data (or did not provide sufficent data, such as including effect estimates but not sample sizes).
Newer reviews were more likely to provide this information than older reviews, but it highlights the importance of data and code sharing as recommended in the PRISMA guidelines [@pagePRISMA2020Statement2020].
When study level data was available, many effects were removed because the pooled sample size was small, or because there were fewer than ten studies on which to perform an Egger's test.
It seems that much of the current screen time research is small in scale, and there is a need for larger, high-quality studies.<span>

@@

Screen time research has a well-established measurement problem, which impacts the individual studies of this umbrella review.
The vast majority of screen time research relies on self-reported data, which not only lacks the nuance required for understanding the effects of screen time, but may also be inaccurate.
In one systematic review on screen time and sleep [@haleScreenTimeSleep2015], 66 of the 67 included studies used self-reported data for *both* the exposure and outcome variable.
It has been established that self-reported screen time data has questionnable validity.
In a meta-analysis of 47 studies comparing self-reported media use with logged measures, Parry et al [@parrySystematicReviewMetaanalysis2021] found that the measures were only moderately correlated ($r = 0.38$), with self-reported problematic usage fairing worse ($r = 0.25$).
Indeed, of 622 studies which measured the screen time of 0---6 year-olds, only 69 provided any sort of psychometric properties for their measure, with only 19 studies reporting validity [@byrneMeasurementScreenTime2021].
While some researchers have started using newer methods of capturing screen behaviours---such as wearable cameras [@smithFeasibilityAutomatedCameras2019] or device-based loggers [@rydingPassiveObjectiveMeasures2020]---these are still not widely adopted.
It may be that the field of screen time research cannot be sufficiently advanced until accurate, validated, and nuanced measures are more widely available and adopted.

>>>>

affiliation:
  - id            : "1"
    institution   : "Institute for Positive Psychology and Education, Australian Catholic University"
  - id            : "2"
    institution   : "School of Health and Behavioural Sciences, Australian Catholic University"
  - id            : "3"
    institution   : "Department of Sport Science and Clinical Biomechanics, University of Southern Denmark"
  - id            : "4"
    institution   : "Centre for Health Research, University of Southern Queensland"
  - id            : "5"
    institution   : "Department of Health Systems and Populations, Faculty of Medicine, Health and Human Sciences, Macquarie University"
  - id            : "6"
    institution   : "School of Kinesiology, Louisiana State University"
  - id            : "7"
    institution   : "School of Medicine and Health, Sydney University"
  - id            : "8"
    institution   : "Health and Wellbeing Centre for Research Innovation, School of Human Movement and Nutrition Sciences, University of Queensland"
  - id            : "9"
    institution   : "Department of Movement and Sport Science, Ghent University"
  - id            : "10"
    institution   : "Institute for Physical Activity and Nutrition, Deakin University"
  - id            : "11"
    institution   : "The Matilda Centre for Research in Mental Health and Substance Use, University of Sydney"
  - id            : "12"
    institution   : "School of Education, Australian Catholic University"

@@

documentclass     : "apa6"
classoption       : "man"
csl               : nature.csl
output            : papaja::apa6_docx
header-includes:
    - \usepackage{makecell}
    - \usepackage{booktabs}
    - \usepackage{colortbl}
---

@@

To facilitate comparisons, we converted effect sizes to Pearson's $r$ using established formulae [@bowmanEffectSizesStatistical2012;@jacobsEstimationBiserialCorrelation2017].
Effect sizes on the original metric are provided in Supplementary File 2.</span>
<span id ="r2_27">Throughout the results section we interpret the size of the effects using Funder and Ozer's guidelines [@funderEvaluatingEffectSize2019]: very small (0.05 < r <= 0.1), small (0.1 < r <= 0.2), medium (0.2 < r <= 0.2), large (0.3 < r <= 0.4), and very large (r >= 0.4).
These are similar to other interpretations based on empirical data [@gignacEffectSizeGuidelines2016].</span>

@@

The most frequently reported exposures were `r manuscript_info$freq_exposures`.
Supplementary File 4 provides a list of all exposures identified.
The most frequently reported outcomes were `r manuscript_info$freq_outcomes`.
<span id ="r2_25">Of the `r manuscript_info$unique_combos` unique exposure/outcome combinations, `r manuscript_info$freq_combos$n1` occurred in only one review, with `r manuscript_info$freq_combos$n2` appearing twice, and `r manuscript_info$freq_combos$ngt2` appearing three or more times.
Full characteristics of the included studies are provided in Table 1.</span>
<span id ="r2_23">After removing reviews with duplicate exposure/outcome combinations, our process yielded `r manuscript_info$unique_effects` unique effect/outcome combinations (retaining multiple effects for different age groups or study designs) contributed from `r manuscript_info$prisma$unique_reviews` reviews.</span>
These effects represent the findings of `r format(manuscript_info$effects_sum$sum_k, big.mark=",")` primary studies comprised of `r format(manuscript_info$effects_sum$sum_n, big.mark=",")` participants.

@@

Screen use research is extensive, varied, and rapidly growing.
Reviews tended to be general (e.g., all screen time) and even when more targeted (e.g., social media) nuances related to specific content (e.g., Instagram vs Facebook) have not been meta-analysed or have not produced credible evidence.
<span id = "r2_29">Fewer than 20\% of the effects identified met our criteria for statistical credibility.
Most studies which did not meet our criteria failed to provide study-level data (or did not provide sufficient data, such as including effect estimates but not sample sizes).
Newer reviews were more likely to provide this information than older reviews, but it highlights the importance of data and code sharing as recommended in the PRISMA guidelines [@pagePRISMA2020Statement2020].
When study level data was available, many effects were removed because the pooled sample size was small, or because there were fewer than ten studies on which to perform an Egger's test.
It seems that much of the current screen time research is small in scale, and there is a need for larger, high-quality studies.<span>

@@

Screen time research has a well-established measurement problem, which impacts the individual studies of this umbrella review.
The vast majority of screen time research relies on self-reported data, which not only lacks the nuance required for understanding the effects of screen time, but may also be inaccurate.
In one systematic review on screen time and sleep [@haleScreenTimeSleep2015], 66 of the 67 included studies used self-reported data for *both* the exposure and outcome variable.
It has been established that self-reported screen time data has questionable validity.
In a meta-analysis of 47 studies comparing self-reported media use with logged measures, Parry et al [@parrySystematicReviewMetaanalysis2021] found that the measures were only moderately correlated ($r = 0.38$), with self-reported problematic usage fairing worse ($r = 0.25$).
Indeed, of 622 studies which measured the screen time of 0---6 year-olds, only 69 provided any sort of psychometric properties for their measure, with only 19 studies reporting validity [@byrneMeasurementScreenTime2021].
While some researchers have started using newer methods of capturing screen behaviours---such as wearable cameras [@smithFeasibilityAutomatedCameras2019] or device-based loggers [@rydingPassiveObjectiveMeasures2020]---these are still not widely adopted.
It may be that the field of screen time research cannot be sufficiently advanced until accurate, validated, and nuanced measures are more widely available and adopted.

### 2023-03-10 reports/revision_letter.Rmd
Some small tweaks

---
title             : "Benefits and risks associated with children's and adolescents' interactions with electronic screens: An umbrella review"
authors           : "Taren Sanders on behalf of co-authors"
journal           : "Nature Human Behaviour"
manuscript        : "NATHUMBEHAV-22071805"
handling_editor   : "Charlotte Payne"

class             : "draft"

output            : revise::letter.pdf
header-includes:
    - \usepackage[utf8x]{inputenc}
---

@@

We have updated the search to September 2022.
We found a substantial amount of new literature had been published since our inital search, with an additional 108 meta-analyses included in the review.
While many of these were updates to existing reviews, we also included 110 new exposure/outcome combinations (or existing combinations for new age groups) that were not found in the original search.
We think that this increase in research demonstrates the level of interest in this area.

@@

We apologise that the quality assessment information was not included in the original submission.
We have added this information to Table 1, along with the information about the sample characteristics requested by the reviewers.
We note that this increases the size of the table, and we have removed it from the text of the manuscript and instead included it as a seperate file.

@@

Please see response to Reviewer 1, \Comment{r1_4}.
We agree that the use of causal langauge from an evidence-base that is mostly cross-sectional would be misleading.
We did not intend to make causal statements, and have reviewed the manuscript for any remaining instances.

@@

Our goal with this review was to provide an overview of the field of screen time research.
A summary of an entire field will, out of necessasity, be more broad that a typical systematic review.
We beleive that there is significant value in providing a single overview to summarise the state of the field.
We have added some additional information to the introduction on this point:

@@

We intended to include the review which captured the most information on the exposure/outcome combination for each effect size.
The assumption was that the largest review would be the most recent, and that he newer reviews would contain the studies located in previous reviews. 
Manual inspection of the reviews which did not contribute any effects suggests that this was sucessful - older reviews tended to only be included if they focused on the general population while a newer review focused on a sub-population.

@@

Note that we do use causal langauge when describing hypothesised mechanisms, and when describing guidelines.
We feel that this is sensible --- guidelines are based on causal assumptions, and while our evidence does not always support them, we are refering to them in the context that they are made (i.e., to instruct parents).

@@

The quote from the *Lancet* editorial is that the understanding is lacking, rather than the evidence.
Our understanding is lacking partly because benefits of screen time (as seen in some educational literature) and reviewed seperately from harms.
Our review aims to address this.
See additional context in the *Introduction*:

@@

`r get_revision("r1_6")`

@@

We had difficulty understanding the reviewers exact concern here.
We are happy to revise this further if the reviewer does not feel that the other changes made to the introduction improve the set of arguements.

@@

We did not *a-priori* choose which outcomes would be included.
Instead, we chose to summarise the literature on for any meta-analysis where screen exposure was the independent variable.
We have removed the word "plausible", as we agree that makes the sentence unclear and the revised phrasing is more aligned with the eligibility critiera.
Our intention in the introduction is not to preempt the results, which means we do not outline the outcomes found in the introduction.

@@

Our apologies for this oversight.
We initially intended to include the quality assessment in Table 1, but felt it made the table too large and moved it to a seperate table.
It seems we inadvertently did not upload the seperate file when submitting the review.
We have now included the quality assessment in Table 1.

@@

2. We were also suprised by the volume of effects which did not meet the criteria for statistical credibility.
We were suprised, for example, that almost a quarter of all effects did not provide study-level data (as might be presented in a forest plot), which we thought would be a low bar to clear.
Almost 40% of those that did have study-level data were derived from samples of less than 1,000 participants (a threshold that has been used by others; 10.1038/s41467-021-24861-8).
And, it seems unwise to place too much emphasis on results where there is an indication of publication bias (i.e., statistically significant Egger's test).
Given that we pre-specified this criteria, we think that it would be inappropriate to lower the standard because the majority of reviews did not reach it.
Instead, we think it reflects the need for the field to raise its standards, especially when it comes to reporting (i.e., making data available, reducing file-drawer problem).

@@

3. We understand the reviewer's concerns regarding the conversions of the effect sizes (although we disagree with the characterisation of our results as "somewhat deceiving").
We chose Pearson's *r* as we wanted to focus not on the precise size of the association, but instead on the *direction* and the *magnitude*.
We think it is worth considering the counter-factual to our approach: had we not converted effect sizes to a common metric readers would likely still make comparisons, but these would be subjective.
We agree that readers should interpret the effect sizes carefully, and have included text to this effect in the *Discussion*:

@@

We are unclear what the reviewer is referring to here, although we think many of the concerns were addressed in \comment{r2_29}.
We feel that a qualitative sytheses of more than 200 effect sizes is outside the scope of this paper.

@@

We did not retrive data from the primary studies when the data were not provided in the paper or by contacting the author.
Our results include more than 3,000 primary studies, and even after accounting for those that provided data in the paper, it would not be feasible to extract all of the remaining data.

@@

Our apologies for this oversight.
Per our response to Reviewer 2 (\comment{r2_15}), we initially intended to include the quality assessment in Table 1.
We removed it because it made the table very large, but did not include the new file with the seperate table.
We have now included the quality assessment in Table 1.

@@

We chose this as most reviews provided mean ages for individual studies but not age ranges, which meant we would not have been able to consistently apply the critiera.

>>>>

---
title             : "Benefits and risks associated with children's and adolescents' interactions with electronic screens: An umbrella review"
authors           : "Taren Sanders on behalf of co-authors"
journal           : "Nature Human Behaviour"
manuscript        : "NATHUMBEHAV-22071805"
handling_editor   : "Charlotte Payne"
link-citations    : true
output            : revise::letter.pdf
header-includes:
    - \usepackage[utf8x]{inputenc}
---

@@

We have updated the search to September 2022.
We found a substantial amount of new literature had been published since our initial search, with an additional 108 meta-analyses included in the review.
While many of these were updates to existing reviews, we also included 110 new exposure/outcome combinations (or existing combinations for new age groups) that were not found in the original search.
We think that this increase in research demonstrates the level of interest in this area.

@@

We apologise that the quality assessment information was not included in the original submission.
We have added this information to Table 1, along with the information about the sample characteristics requested by the reviewers.
We note that this increases the size of the table, and we have removed it from the text of the manuscript and instead included it as a separate file.

@@

Please see response to Reviewer 1, \Comment{r1_4}.
We agree that the use of causal language from an evidence-base that is mostly cross-sectional would be misleading.
We did not intend to make causal statements, and have reviewed the manuscript for any remaining instances.

@@

Our goal with this review was to provide an overview of the field of screen time research.
A summary of an entire field will, out of necessity, be more broad that a typical systematic review.
We believe that there is significant value in providing a single overview to summarise the state of the field.
We have added some additional information to the introduction on this point:

@@

We intended to include the review which captured the most information on the exposure/outcome combination for each effect size.
The assumption was that the largest review would be the most recent, and that he newer reviews would contain the studies located in previous reviews. 
Manual inspection of the reviews which did not contribute any effects suggests that this was successful - older reviews tended to only be included if they focused on the general population while a newer review focused on a sub-population.

@@

Note that we do use causal language when describing hypothesised mechanisms, and when describing guidelines.
We feel that this is sensible --- guidelines are based on causal assumptions, and while our evidence does not always support them, we are referring to them in the context that they are made (i.e., to instruct parents).

@@

The quote from the *Lancet* editorial is that the understanding is lacking, rather than the evidence.
Our understanding is lacking partly because benefits of screen time (as seen in some educational literature) and reviewed separately from harms.
Our review aims to address this.
See additional context in the *Introduction*:

@@

`r get_revision("r2_6")`

@@

We had difficulty understanding the reviewers exact concern here.
We are happy to revise this further if the reviewer does not feel that the other changes made to the introduction improve the set of arguments.

@@

We did not *a-priori* choose which outcomes would be included.
Instead, we chose to summarise the literature on for any meta-analysis where screen exposure was the independent variable.
We have removed the word "plausible", as we agree that makes the sentence unclear and the revised phrasing is more aligned with the eligibility criteria.
Our intention in the introduction is not to preempt the results, which means we do not outline the outcomes found in the introduction.

@@

Our apologies for this oversight.
We initially intended to include the quality assessment in Table 1, but felt it made the table too large and moved it to a separate table.
It seems we inadvertently did not upload the separate file when submitting the review.
We have now included the quality assessment in Table 1.

@@

2. We were also surprised by the volume of effects which did not meet the criteria for statistical credibility.
We were surprised, for example, that almost a quarter of all effects did not provide study-level data (as might be presented in a forest plot), which we thought would be a low bar to clear.
Almost 40% of those that did have study-level data were derived from samples of less than 1,000 participants (a threshold that has been used by others; 10.1038/s41467-021-24861-8).
And, it seems unwise to place too much emphasis on results where there is an indication of publication bias (i.e., statistically significant Egger's test).
Given that we pre-specified this criteria, we think that it would be inappropriate to lower the standard because the majority of reviews did not reach it.
Instead, we think it reflects the need for the field to raise its standards, especially when it comes to reporting (i.e., making data available, reducing file-drawer problem).

@@

3. We understand the reviewer's concerns regarding the conversions of the effect sizes (although we disagree with the characterisation of our results as "somewhat deceiving").
We chose Pearson's *r* as we wanted to focus not on the precise size of the association, but instead on the *direction* and the *magnitude*.
We think it is worth considering the counterfactual to our approach: had we not converted effect sizes to a common metric readers would likely still make comparisons, but these would be subjective.
We agree that readers should interpret the effect sizes carefully, and have included text to this effect in the *Discussion*:

@@

We are unclear what the reviewer is referring to here, although we think many of the concerns were addressed in \comment{r2_29}.
We feel that a qualitative synthesis of more than 200 effect sizes is outside the scope of this paper.

@@

We did not retrieve data from the primary studies when the data were not provided in the paper or by contacting the author.
Our results include more than 3,000 primary studies, and even after accounting for those that provided data in the paper, it would not be feasible to extract all of the remaining data.

@@

Our apologies for this oversight.
Per our response to Reviewer 2 (\comment{r2_15}), we initially intended to include the quality assessment in Table 1.
We removed it because it made the table very large, but did not include the new file with the separate table.
We have now included the quality assessment in Table 1.

@@

We chose this as most reviews provided mean ages for individual studies but not age ranges, which meant we would not have been able to consistently apply the criteria.
